\chapter*{Acknowledgements}

I want to extend my deepest gratitude to my final year project supervisor, \hyperlink{https://www.gla.ac.uk/schools/computing/staff/angelosmarnerides/}{Dr. Angelos K. Marnerides}, for the continual guidance, advice, and support that I have received throughout my project. Throughout Angelos' consultation, I have learned significantly from his expertise and research contributions. The time spent working with him has been both a pleasure and an honour.

The development of VisiBot would not have been possible without the contributions of \hyperlink{https://badpackets.net/}{Bad Packets}. I want to offer my deepest gratitude towards both the secretary and director of Bad Packets, Troy Mursch and Mathew Woodyard, for providing full access to the Bad Packets Cyber Threat Intelligence service and hosting platforms, offering excellent advice and support throughout my project. I would also like to thank Polina Stolpovskaya and associates at \hyperlink{https://ipinfo.io/}{IPInfo} for providing access to their various IP address API services and the support offered throughout my academic research. Additionally, thanks should also go to \hyperlink{https://www.virustotal.com/gui/}{VirusTotal} for providing special access to premium API features in support for my research, and \hyperlink{https://www.maxmind.com}{MaxMind} for providing access to their GeoIP2 databases service free of charge. I want to offer many thanks to \hyperlink{https://github.com/danieluhricek}{Daniel Uhříček}, developer of the open-source \hyperlink{https://github.com/danieluhricek/LiSa}{LiSa sandbox} for his hard work and extensive support. 

Finally, I am incredibly grateful for the support offered by my family, friends, and fellow students at the University of Glasgow throughout my final year of studies. I want to close my acknowledgements by thanking the staff at the University of Glasgow School of Computing Science, and more so the head of honours,  \hyperlink{https://www.gla.ac.uk/schools/computing/staff/gethinnorman/}{Dr. Gethin Norman}, for providing unparalleled support throughout the COVID-19 pandemic. 